---
title: "Pedagogy 7/6/17 - purrr"
author: "Caleb Scheidel"
date: "6/29/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(purrr)
library(devtools)
library(repurrrsive)
library(listviewer)
library(knitr)
library(tibble)
library(tidyverse)

```

# Tutorial Outline

1. Introduce advantages of `purrr` package
2. Use GoT dataset to run through examples of commonly used `purrr` functions: `map()`, `map2()`, and `pmap()`
3. Walk through a real example where I learned the hard way that `purrr` is appropriate for use in a Methods project (Note: not in this document)
 
## GoT Characters Dataset

For this tutorial, we will be using the `got_chars` dataset from the `repurrrsive` package.  It contains data on all of the point-of-view characters from the Game of Thrones book series in a list of lists format.  The highest-level list contains 29 characters, with each character having a list of 18 variables containing information pertaining to each character, i.e. `name`, `gender`, and `culture`.  For this tutorial, we will be using the `got_chars` dataset from the `repurrrsive` package.  It contains data on all of the point-of-view characters from the Game of Thrones book series in a list of lists format.  The highest level list contains 29 characters, with each character having a list of 18 variables containing information pertaining to each character, i.e. `name`, `gender`, and `culture`.

```{r}

#Only viewing elements of the first character's list
str(got_chars[1])

```

## Introduction to map(): extract elements

`map()` is a function for applying a function to each element of a list.

template for basic map usage: `map(YOUR_LIST, YOUR_FUNCTION)`

First, a quick example, taking the square root of a list of three numbers:

```{r}

# Quick example
map(c(9, 16, 26), sqrt)

```

Using the GoT characters dataset, pulling out the names from specified characters:

```{r}

#name and position shortcuts
map(got_chars[1:4], "name")
map(got_chars[5:8], 3) #name is in third position

#can rewrite using the pipe
got_chars[1:4] %>% 
  map("name")

```


### Type-specific map

You can specify type-specific variants to map, such as `map_chr()`, `map_lgl()`, `map_int()`, and `map_dbl()`:

```{r}

map_chr(got_chars[9:12], "name")

map_int(got_chars[9:12], "id")

map_lgl(got_chars[9:12], "alive")

```

### Extract multiple values

Recall map() usage: `map(.x, .f, ...)`

The function .f will be `[`.  We use `...` to pass the character vector of the names of desired elements 

```{r warning=FALSE, message=FALSE}

#how to retrieve multiple elements without using map()
got_chars[[3]][c("name", "culture", "gender", "born")]

x <- map(got_chars, `[`, c("name", "culture", "gender", "born"))
str(x[16:17])

#This can be rewritten using the pipe and extract() function from magrittr:
library(magrittr)

x <- got_chars %>% 
  map(extract,  c("name", "culture", "gender", "born"))
  
str(x[18:19])

```

To get output into a dataframe, use `map_df()`:

```{r}

got_char <- map_df(got_chars, extract, c("name", "culture", "gender", "id", "born", "alive"))

got_char %>%
  kable()

```

## Specifying the function in map() + parallel mapping

### map() function specification

Recall map() usage: `map(.x, .f, ...)`

We demonstrate three more ways to specify a general function, `.f`:

- an existing function
- an anonymous function, defined on the fly
- a formula

We will continue to work with the GoT character list, `got_chars`.  Each character can have aliases, which are stored as a list for each character.  We pull out the aliases for three characters to use in this demo:

```{r}

aliases <- set_names(map(got_chars, "aliases"),
                     map_chr(got_chars, "name"))

(aliases <- aliases[c("Tyrion Lannister", "Daenerys Targaryen", "Jon Snow")])

```

We want to present each character's aliases combined together in one variable, separated by " | ".  There are three ways to go about doing this using `map()`

### Existing function

We can define our own function and use it in `map()`:

```{r}

my_fun <- function(x) paste(x, collapse = " | ")
map(aliases, my_fun)

```

### Anonymous function, conventional

We can also define the function on the fly in the `map()` function itself:

```{r}

map(aliases, function(x) paste(x, collapse = " | "))

#or rewrite it as
map(aliases, paste, collapse = " | ")

```

### Anonymous function, formula

purrr provides a concise way to define an anonymous function: as a formula.  This should start with the `~` symbol and then look like a typical expression.  Use `.x` to refer to the input, i.e. an individual element of the primary vector or list.

```{r}

map(aliases, ~ paste(.x, collapse = " | "))

```

#### Development process

It's rare to write these from scratch and work well the first time.  A typical development process would look like this:

```{r}

(a <- map(got_chars, "aliases")[[19]]) #OOPS! empty, so useless

(a <- map(got_chars, "aliases")[[16]]) #ok good

paste(a, collapse = " | ") #ok good

#Now we know
got_chars[15:17] %>% 
  map("aliases") %>% 
  map_chr(paste, collapse = " | ")

```

### List to data frame

Since we've simplified the aliases to a single string for each character, we can hold them as an atomic character vector instead of as a list.  The `tibble::enframe()` function takes a named vector and promotes the names to a proper variable.

```{r}

aliases <- set_names(map(got_chars, "aliases"), map_chr(got_chars, "name"))

aliases_df <- map_chr(aliases[c(3, 10, 20, 24)], ~ paste(.x, collapse = " | ")) %>% 
  tibble::enframe(value = "aliases")

aliases_df %>% 
  head() %>% 
  kable()

```

### Parallel map using `map2()`

You can use `map2()` to map a function over two vectors or lists in parallel.  Here is the usage:

- `map2(.x, .y, .f, ...)` 

or

- `map2(INPUT_ONE, INPUT_TWO, FUNCTION_TO_APPLY, OPTIONAL_OTHER_STUFF)`

It can also be type specific: `map2_chr()`, `map2_dbl()`, etc.

For our example, we will paste the character's name and the date and location of his or her birth to get a sentence.

First, obtain the two inputs:

```{r}

nms <- got_chars %>% 
  map_chr("name")

birth <- got_chars %>% 
  map_chr("born")

```

Map over both with an existing function, defined by us:

```{r}

my_fun <- function(x, y) paste(x, "was born", y)
map2_chr(nms, birth, my_fun) %>% 
  head()

```

Or with an anonymous function, conventional form:

```{r}

map2_chr(nms, birth, function(x, y) paste(x, "was born", y)) %>% head()

```

Or with an anonymous function via formula.  Use `.x` and `.y` to refer to individual elements of the two primary inputs:

```{r}

map2_chr(nms[16:18], birth[16:18], ~ paste(.x, "was born", .y))

```

### Parallel mapping using `pmap()`

Use `pmap()` to map a function over two or more vectors or lists in parallel.

Usage:

- `pmap(.l, .f, ...)`

or

- `pmap(LIST_OF_INPUT_LISTS, FUNCTION_TO_APPLY, OPTIONAL_OTHER_STUFF)`

Example with a function that creates a sentence for each character stating their name, the number of aliases, and allegiances:

```{r}

df <- got_chars %>% {
  tibble::tibble(
    name = map_chr(., "name"),
    aliases = map(., "aliases"),
    allegiances = map(., "allegiances")
  )
}

my_fun <- function(name, aliases, allegiances) {
  paste(name, "has", length(aliases), "aliases and", length(allegiances), "allegiances")
}

df %>% 
  pmap_chr(my_fun) %>% 
  tail()

```
